Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-07-22 Paper/Poster/poster.tex
getting there, some feedback from Ron included

\pbox{0.957\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{

%%% Unisiegel
\begin{minipage}[c][10cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps} \\
    \vspace{.5cm}
    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
  \end{center}
\end{minipage}
%\begin{minipage}[c][9cm][c]{0.1\textwidth}
%  \begin{center}
%    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
%  \end{center}
%\end{minipage}
%%% Title
\begin{minipage}[c][9cm][c]{0.58\textwidth}
  \begin{center}
    {\sc \Huge Clustering Beat-Chroma Patterns}\\[5mm]
  {\sc \Huge in a Large Music Database}\\[10mm]
    {\Large Thierry Bertin-Mahieux $^\dagger$, Ron J. Weiss $^\ddagger$ and Daniel P. W. Ellis $^\dagger$\\[3.5mm]
     \large $\dagger$ LabROSA, Dept. of Electrical Engineering, Columbia University \\
     \large $\ddagger$ MARL, New York University \\
    \tt{\{thierry,ronw,dpwe\}@ee.columbia.edu}}
  \end{center}
\end{minipage}
%%% GK-Logo
\begin{minipage}[c][7cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=10cm,angle=0]{MARL_logo_high_res.eps} \\
    \vspace{1.8cm}
    \includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
  \end{center}
\end{minipage}

@@

%%% Section 1
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Introduction
    \end{center}}
\end{center}

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Availability of very large collections of music audio:
can we infer anything about the underlying structure and common
features of e.g. commercial pop music?
\item Our interest: tonal content of the music -- i.e. the
harmony and melody.
\item Beat-synchronous chromagrams: rich enough
to generate musically-relevant results, simplified enough to
abstract away instrumentation and other stylistic details.
\end{list}

\vspace{-.3cm}
\makebox[\columnwidth]{ % replace framebox by makebox
\hspace{-4cm}
\begin{minipage}[l]{.78\columnwidth}
\myfig[0]{code_axis.ps}{.45}
\end{minipage}
\hspace{-3.5cm}
\begin{minipage}[l]{.4\columnwidth}
Typical learned codeword spanning $2$ bars, each bar
considered to be $4$ beats. Codebook size was $200$.\\

@@

%%% Section 2
\vspace{1.5cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Audio Features - Echo Nest
    \end{center}}
\end{center}

@@

\textbf{Chromagrams from the Echo Nest online API.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
\item For any song, EchoNest provides a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars. Beats may span or 
subdivide segments; bars span multiple beats.
\item Averaging per-segment chroma over beat times results in a
beat-synchronous chroma feature representation.
\item Dataset size: $43,000$ songs.
\end{list}

\vspace{.5cm}
\textbf{Beat-Chroma Patches.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item We use Echo Nest analysis to break a song into a collection of 
beat-chroma ``patches'', typically one or two bars in length.
\item $82\%$ of the bars in our data were 4 beats long.
\item Other cases: we resample patches to a fixed length of 4 beats.
\item We rotate patches so that the first row contains the most energy.
Each patch is normalized independently.
\end{list}

@@

%%% Section 3
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Vector Quantization
    \end{center}}
\end{center}

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Vector Quantization algorithm \cite{Gersho1991} to cluster beat-chroma
patches
\item VQ can be seen as online $k$-means
\item VQ initialized with $k$ random patches from the data
\item VQ, although not optimal, scales linearly with
the number of patches seen
\end{list}

@@

%%% Section 4
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Pattern Analysis
    \end{center}}
\end{center}

@@

%%% Section 5
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Experiments
    \end{center}}
\end{center}

@@

\textbf{Artist recognition task}.
We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres.  
Previously published results using GMMs on MFCC features achieve an 
accuracy of $59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$  \cite{Ellis2007}.

We get an accuracy of $\mathbf{23.4\%}$, random baseline 
is around $5\%$. 
The confusion matrix is shown here, note 
that certain artists are recognized at an accuracy far above the average.

@@

\begin{minipage}[r]{.5\columnwidth}
\textbf{Bar alignment task}.
Since the clustering described is based
on the segmentation of the signal in to bars, 
the codewords should contain information related to bar
alignment, such as the presence of a strong beat on the first beat.
It is the case $62\%$ of the times, see paper for details.
\end{minipage}

@@

%%% Section 6
\vspace{0cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \large Conclusion
    \end{center}}
\end{center}

@@

%%% References
\begin{small}
\bibliographystyle{IEEEtran}
\bibliography{tbm_bib}
\end{small}

% authors
\vspace{-2cm}
\myfig[0]{authors.eps}{.8}
% figure caption
\vspace{-2cm}
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{center}
\begin{small}
Looking forward to discuss our work with you!
\end{small}
\end{center}
\end{minipage}
\end{center}

>>>>

\makebox[0.957\textwidth]{

%%% Unisiegel
\begin{minipage}[c][7cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=7cm,angle=0]{labrosa-logo.eps} \\
    \vspace{.5cm}
    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
  \end{center}
\end{minipage}
%\begin{minipage}[c][9cm][c]{0.1\textwidth}
%  \begin{center}
%    \includegraphics[width=7cm,angle=0]{seas-logo.eps}
%  \end{center}
%\end{minipage}

\pbox{0.57\textwidth}{}{linewidth=2mm,framearc=0.3,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
%%% Title
\begin{center}
\begin{minipage}[c][10cm][c]{0.90\textwidth}
  \begin{center}
    {\sc \Huge Clustering Beat-Chroma Patterns}\\[5mm]
  {\sc \Huge in a Large Music Database}\\[10mm]
    {\Large Thierry Bertin-Mahieux $^\dagger$, Ron J. Weiss $^\ddagger$ and Daniel P. W. Ellis $^\dagger$\\[3.5mm]
     \large $\dagger$ LabROSA, Dept. of Electrical Engineering, Columbia University \\
     \large $\ddagger$ Music and Audio Research Laboratory, New York University \\
    \tt{\{thierry,ronw,dpwe\}@ee.columbia.edu}}
  \end{center}
\end{minipage}
\end{center}
}

%%% GK-Logo
\begin{minipage}[c][7cm][c]{0.2\textwidth}
  \begin{center}
    \includegraphics[width=10cm,angle=0]{MARL_logo_high_res.eps} \\
    \vspace{1.8cm}
    %\includegraphics[width=8cm,angle=0]{nyu_music_tech_logo_2_2_bw.eps}
    \includegraphics[width=14cm,angle=0]{NYU_logo.eps}
  \end{center}
\end{minipage}

@@

%%% Section 1
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Introduction
    \end{center}}
\end{center}

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Can we infer anything about the underlying structure and common
features of commercial pop music by analyzing very large collections
of music audio?
\item Our interest: tonal content of the music -- i.e. the
harmony and melody.
\item Beat-synchronous chromagrams: rich enough
to generate musically-relevant results, simplified enough to
abstract away instrumentation and other stylistic details.
\end{list}

\vspace{-.3cm}
\makebox[\columnwidth]{ % replace framebox by makebox
\hspace{-4cm}
\begin{minipage}[l]{.78\columnwidth}
\myfig[0]{code_axis.ps}{.40}
\end{minipage}
\hspace{-3.5cm}
\begin{minipage}[l]{.4\columnwidth}
Typical %learned 
codeword spanning $2$ bars, each bar
considered to be $4$ beats. Codebook size was $200$.\\

@@

%%% Section 2
\vspace{1.5cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Audio Features - Echo Nest
    \end{center}}
\end{center}

@@

\textbf{Chromagrams from the Echo Nest online API.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Feature analysis based on Echo Nest analyze API \cite{EchoNest}.
\item For any song, EchoNest provides a chroma
vector (length $12$) for every music event (called ``segment''), and a
segmentation of the song into beats and bars.
\item Averaging per-segment chroma over beat times results in a
beat-synchronous chroma feature representation.
\item Dataset size: $43,000$ songs.
\end{list}

\vspace{.5cm}
\textbf{Beat-Chroma Patches.}
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item We use Echo Nest analysis to break a song into a collection of 
beat-chroma ``patches'', typically one or two bars in length.
\item $82\%$ of the bars in our data were 4 beats long.
\item Other cases: we resample patches to a fixed length of 4 beats.
\item We rotate patches so that the first row contains the most energy.
This is analogous to key normalization, but
each patch is normalized independently.
\end{list}

@@

%%% Section 3
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Vector Quantization
    \end{center}}
\end{center}

@@

\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Vector Quantization algorithm \cite{Gersho1991} to cluster beat-chroma
patches
\item Can be seen as online $k$-means
\item Initialized with $k$ random patches from the data
\item Although not optimal, scales linearly with
the number of patches seen
\end{list}

@@

%%% Section 4
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Pattern Analysis
    \end{center}}
\end{center}

@@

%%% Section 5
\vspace{2cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Experiments
    \end{center}}
\end{center}

@@

\textbf{Artist recognition task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item {\it artist20} data set: $1402$ songs from $20$ artists.
\item Baseline: GMM on MFCC ($59\%$), GMM on chromas ($33\%$),
random ($5\%$).
\item We get an accuracy of $\mathbf{23.4\%}$.
\end{list}
%We use the {\it artist20} data set: $1402$ songs from $20$ artists, 
%mostly rock and pop of different subgenres.  
%Previously published results using GMMs on MFCC features achieve an 
%accuracy of $59\%$, whereas using only chroma as a representation yields an
%accuracy of $33\%$  \cite{Ellis2007}.
%
%We get an accuracy of $\mathbf{23.4\%}$, random baseline 
%is around $5\%$. 
%The confusion matrix is shown here, note 
%that certain artists are recognized at an accuracy far above the average.

@@

\begin{minipage}[r]{.5\columnwidth}
\textbf{Bar alignment task}.
\begin{list}{\labelitemi}{\leftmargin=2em \labelsep=1em}
\item Clustering based on the signal segmented in bars.
\item Codewords should contain bar information (e.g. a strong
first beat).
\item Our encoding find the right alignment $62\%$ of the time.
\end{list}
%Since the clustering described is based
%on the segmentation of the signal in to bars, 
%the codewords should contain information related to bar
%alignment, such as the presence of a strong beat on the first beat.
%It is the case $62\%$ of the times, see paper for details.
\end{minipage}

@@

%%% Section 6
\vspace{0cm}
\begin{center}
  \pbox{0.8\columnwidth}{}{linewidth=2mm,framearc=0.1,linecolor=lightred,fillstyle=gradient,gradangle=0,gradbegin=white,gradend=whitepink,gradmidpoint=1.0,framesep=1em}{
    \begin{center}
      \LARGE Conclusion
    \end{center}}
\end{center}

@@

%%% References
%\begin{small}
\bibliographystyle{IEEEtran}
\bibliography{tbm_bib}
%\end{small}

\iffalse
% authors
\vspace{-2cm}
\myfig[0]{authors.eps}{.8}
% figure caption
\vspace{-2cm}
\begin{center}
\begin{minipage}[c]{.7\columnwidth}
\begin{center}
\begin{small}
Looking forward to discuss our work with you!
\end{small}
\end{center}
\end{minipage}
\end{center}
\fi
